### 2014-09-08 writeups/context-dependence.tex
computing transition matrix across a general tree does not error (but not right yet? see tree-test.R)

>>>>

\subsection*{Algorithm}

Fix one tip of the tree, $a$, to be the taxa where we count ``long'' sequences, let $\rho$ be the root, and let $b_1, \ldots, b_n$ be the remaining tips,
where we observe ``short'' sequences.
Let $v_0=\rho, v_1, \ldots, v_\ell = a$ be the path from the root down to $a$,
let $u_1, u_2, \ldots, u_m$ be the remaining internal nodes,
and for each $v_i$ let $u(v_i)$ be the offspring of $v_i$ that is not one of the other $v$.
For each node in the tree,
we will compute the matrix of probabilities that a ``long'' sequence observed at that node
is matched by the ``short'' sequences seen on all the tips of the tree
further away from $a$ (i.e.\ below that node if it was rooted at $a$).
For a node $w$ this will be stored as a $|\calS|^{m+2\ell} \times |\calS|^{cm}$ matrix, for some $c$,
and denoted $M(w)$.

  \item Set $M(b_i) = P$ for each $b_i$.

  \item While there are $u_j$ whose children $w_1$ and $w_2$ both have matrices computed,
    \begin{enumerate}
      \item let
        \begin{align}
          M(u_j)_{x,(y,z)} = \left( e^{t(w_1) G} M(w_1) \right)_{x,y} \left( e^{t(w_2) G} M(w_2)_{x,z} \right)_{x,z} ,
        \end{align}
        where $t(w_1)$ and $t(w_2)$ are the lengths of the respective branches.
    \end{enumerate}

  \item Let 
    \begin{align}
      M(\rho)_{x,y} = \pi(x) \left( e^{t(w(\rho)) G} M(w(\rho)) \right)_{x,y}
    \end{align}

  \item For each $1 \le k < \ell$,
    \begin{enumerate}
      \item let
        \begin{align}
          M(v_k)_{x,(y,z)} = \left( e^{t(v_k) G} M(v_{k-1}) \right)_{x,y} \left( e^{t(u(v_k)) G} M(u(v_{k-1})) \right)_{x,z}  .
        \end{align}
    \end{enumerate}

  \item Finally, let
    \begin{align}
      M(a)_{x,y} = \left( e^{t(a) G} M(v_{\ell-1}) \right)_{x,y}  .
    \end{align}
